---
title: "Parallel Computing"
output: html_notebook
thanks: "Replication files are available on the author's Github account."
author:
- name: Felix Taschbach
  affiliation: Maastricht University
---

## Load Packages
We begin by installing and loading any required packages. The following code first installs the package pacman, if necessary, and then uses pacman to install, if required, and load the packages needed to run this script.
```{r results="hide"}
# Clear all variables 
rm(list=ls()) 
options(java.parameters="-Xmx16g")

# install pacman if necessary
if (!require("pacman")) suppressPackageStartupMessages(install.packages("pacman"))
# install required packages
pacman::p_load("parallel","rcdk","itertools","doParallel","multidplyr")
# install.packages("devtools")
#devtools::install_github("hadley/multidplyr")

```
## Import the data
Now, I load the `aid624202.sdf` file and compute the descriptor matrix. As the file is over 1 gb, it is imported using an iterator. The iterator goes through every molecule one by one and adds it to the `mols` list, either until all molecules are imported, or until a pre-set number of molecules is imported.
```{r}
# the total amount of molecules in the sdf file
max_n = 377550
# the amount of molecules that will be imported
n = 1000

# iload.molecules requires the full path
# creates an iterator to import the `aid624202.sdf` file
iter <- iload.molecules('/Users/felix/Google Drive/Uni/Systems Biology/Year 2/Scientific Programming/scientificProgramming2/aid624202.sdf', type='sdf')
# creates a txtProgressBar object to display the progress of importing the selected molecules
pb <- txtProgressBar(min = 0, max = n, style = 3)
counter = 0   # initiate counter
mols = list()   # initiate list to hold the selected molecules
while(hasNext(iter)) {  # runs until 
  counter = counter + 1 # update the counter
  mols[[counter]] = nextElem(iter)   # save the selected molecule in the list
  setTxtProgressBar(pb, counter)
  if(counter == n){
    break
  }
}
close(pb)
```
## Calculate the molecular descriptors in parallel
Now, I distribute the data among the cores and compute the molecular descriptors of each imported molecule. The code is set up so that it can go parallel over the molecules or over the descriptors. As we are only interested in the amount of time it takes to calculate the descriptors and not in the descriptors themselves, only the seconds elapsed will be saved.
```{r}
# detects the amount of cores of the computer running this script
numCores <- detectCores() - 1
# get a list of all descriptors that will be calculated
descNames <- unique(unlist(sapply(get.desc.categories()[2:4], get.desc.names)))

pb <- txtProgressBar(min = 0, max = numCores, style = 3)

# convert the molecules into smiles
smiles = lapply(unlist(mols),get.smiles)
# initialise list that will hold the amount of seconds elapsed during each calculation
time = list()
# set the initial time without any parallel computing
#time[0] = system.time(foreach(i=1:n,.combine=rbind) %do% eval.desc(parse.smiles(smiles[[i]]),descNames,verbose=F))[3]
#time[0] = system.time(foreach(i=1:length(descNames),.combine=cbind) %dopar% eval.desc(parse.smles(unlist(similes)),descNames[i],verbose=F))[3]

setTxtProgressBar(pb, 0)

# returns the seconds elapsed during molecular descriptor calculation for various amounts of cores
for(ncores in 1:numCores){
  cl <- makeCluster(ncores)     # create cluster
  clusterExport(cl,"descNames") # export descriptor names to each core
  clusterExport(cl,"smiles")    # export the molecules as smiles to each core
  clusterEvalQ(cl, pacman::p_load("parallel","rcdk","itertools","doParallel")) # load required packages in each core
  
  registerDoParallel(cl) # make the next foreach call parallel
  time[ncores] = system.time(foreach(i=1:n,.combine=rbind) %dopar% eval.desc(parse.smiles(smiles[[i]]),descNames,verbose=F))[3]
#time[ncores]=system.time(foreach(i=1:length(descNames),.combine=cbind) %dopar% eval.desc(parse.smiles(unlist(smiles)),descNames[i],verbose=F))[3]
  setTxtProgressBar(pb, ncores)
  stopCluster(cl)
}
```
## Finally, I plot the results.
```{r}
plot(unlist(time),xlab="amount of cores used",ylab="time in seconds",main=paste0("Time it took to calculate descriptors for ",n," molecules"))
```



